---
title: "Contamination Experiment"
output:
  html_notebook: default
  pdf_document: default
---

We conducted a contamination experiment with a blend of GLVs.  We sampled with both DCSE and DHS with either 1) just the GLV blend, 2) just tea plants, or 3) tea plants with the blend of GLVs at the soil level.

Unfortunately background readings are missing for two treatment * method runs (one because of low signal, one because we ran out of tenax tubes).  I'm going to use averages from the other two reps for these.  Not sure if that is best practice yet.

```{r}
library(tidyverse)
library(xtable)
```

```{r}
input <- read_csv("Contamination.csv")[ , 1:5]
input
background <- read_csv("Contamination background.csv")
background %>% group_by(Method, Treatment, Compound) %>% 
  mutate(mean.conc = mean(Concentration, na.rm = TRUE)) -> mean.background

#replace NAs with means from other two reps for that experiment
mean.background$Concentration[is.na(mean.background$Concentration)] <- mean.background$mean.conc[is.na(mean.background$Concentration)]

data <- input
data$Concentration = input$Concentration - mean.background$Concentration
data$Concentration[data$Concentration < 0] <- 0
data
```
For a paired t-test need both x and y.  So need to wrangle to have separate vectors for each compound x method combination for just plants and plants + glvs (both)

```{r}
data %>% filter(Method == "DCSE") %>% 
  filter(Treatment != "glvs") %>% 
  spread(key = Compound, value = Concentration) -> DCSE.data
DCSE.data

data %>% filter(Method == "DHS") %>% 
  filter(Treatment != "glvs") %>% 
  spread(key = Compound, value = Concentration) -> DHS.data
DHS.data
```
Ok, now I need to do t-tests for each compound
```{r}
fnx <- function(df){
  results <- data.frame(Compound = colnames(df)[4:9], p.value = rep(NA,6))

  for(i in 4:9){
    x <- df[1:3, i][[1]]
    y <- df[4:6, i][[1]]
    result <- t.test(x = x, y = y, paired = TRUE)
    results$statistic[i-3] <- result$statistic
    results$df[i-3] <- result$parameter
    results$p.value[i-3]<- result$p.value
  }
  return(results)
}

fnx(DCSE.data)
fnx(DHS.data)
```
Cool, now make a fancy table with `xtable`
```{r}
#DCSE data first
DCSE.data %>%  gather(key = Compound, value = Concentration, -Treatment, -Method, -Plant) -> summary1
summary1
both.data <- filter(summary1, Treatment == "both")[,-c(1:2)]
plants.data <- filter(summary1, Treatment == "plants")[, -c(1:2)]

summary2 <- data.frame(Plant = both.data$Plant, Compound = both.data$Compound, Just.Plant = plants.data$Concentration, GLV.and.Plant = both.data$Concentration)
summary2$Difference <- summary2$GLV.and.Plant - summary2$Just.Plant
summary2 %>% 
  group_by(Compound) %>% 
  summarise_all(funs(mean)) %>% 
  select(-Plant) %>% 
  left_join(fnx(DCSE.data)) -> summary3
summary3$p.value[is.nan(summary3$p.value)] <- NA
colnames(summary3) <- c("Compound", "Plant Only", "Plant + GLV blend", "Difference", "p-Value")
write_csv(summary3, "DCSE contamination output.csv")


DHS.data %>%  gather(key = Compound, value = Concentration, -Treatment, -Method, -Plant) -> summary1.DHS
summary1.DHS
both.data.DHS <- filter(summary1.DHS, Treatment == "both")[,-c(1:2)]
plants.data.DHS <- filter(summary1.DHS, Treatment == "plants")[, -c(1:2)]

summary2.DHS <- data.frame(Plant = both.data.DHS$Plant, Compound = both.data.DHS$Compound, Just.Plant = plants.data.DHS$Concentration, GLV.and.Plant = both.data.DHS$Concentration)
summary2.DHS$Difference <- summary2.DHS$GLV.and.Plant - summary2.DHS$Just.Plant

summary2.DHS %>% 
  group_by(Compound) %>% 
  summarise_all(funs(mean)) %>% 
  select(-Plant) %>% 
  left_join(fnx(DHS.data)) -> summary3.DHS

colnames(summary3.DHS) <- c("Compound", "Plant Only", "Plant + GLV blend", "Difference", "p-Value")
write_csv(summary3.DHS, "DHS contamination output.csv")

```




```{r}
output <- xtable(summary3, digits = c(0,0,2,2,2,3))
```

```{r, results='asis'}
print(output, NA.string = "NA", type = "html")
```
