---
title: "DHS comparison"
output: html_notebook
---

I copied this over from a .R file on Nov 17, 2016 and simplified some of the code. Since we're only interested in DHS by tenax and DCSE by twister, I've deleted quite a bit of code.

```{r, include=FALSE}
#setwd("/Users/scottericr/Google Drive/R Code/Twister Data")
library(reshape2)
library(tidyr)
library(dplyr)
library(venn)
library(ggplot2)
library(readr)
#library(rJava)
#library(xlsx)
source("~/Google Drive/R Code/myfunctions.R")
```
#Importing the data

```{r}
#input <- read_csv("DHS Comparison updated 4-7-2017.csv") #Nicole added back two compounds she previously removed as contaminants.  They're not actually contaminants.
input <- read_csv("DHS Comparison updated 6-4-2017.csv") #Nicole updated to remove some compounds that are plasticizers and others that are below a modified LOQ

input2 <- input[1:12, ]
#glimpse(input2)

input2[is.na(input2)] <- 0  #NA's represent 0's


#Read in annotated library for later extraction of LogKOW and vapor pressure
#library <- read_csv("annotated library mar 13 2017.csv")
#library <- read_csv("annotated library apr 6 2017.csv")
```

create treatment variables
```{r}
vars <- colsplit(input2$Treatment, "\\.", c("Method", "Treatment", "Rep"))
vars

len <- dim(input2)[2]
data <- cbind(vars$Method, vars$Treatment, input2[ , 2:len])
colnames(data)[1:2] <- c("Method","Treatment")
head(data)
#glimpse(data)
```

find any compounds that are constants
```{r}
counts <- head(summarise_each(data, funs(n_distinct)))
rm <- which(counts <= 1)

if(length(rm) != 0){
data <- subset(data, select = -rm)}
#glimpse(data)
```

## With only DHS Tenax vs. DCSE Twister
```{r}
data %>% group_by(Method) %>% summarise_all(funs(mean)) -> compare.data
compare.data[ , -(1:2)] %>% t() %>% as.data.frame() -> compare.data
colnames(compare.data) <- c("DHS", "DCSE")
compare.data$Compound <- rownames(compare.data)
rownames(compare.data) <- NULL

##Venn Diagram
method.logical <- compare.data
method.logical$DHS[method.logical$DHS != 0] <- 1
method.logical$DCSE[method.logical$DCSE !=0] <- 1

venn(method.logical[ , 1:2], cexil = 1, cexsn = .85)

#Extract Unique Compounds
select(compare.data, Compound, everything()) %>% filter(DHS == 0 & DCSE != 0 | DHS !=0 & DCSE ==0) -> method_unique_compounds
#left_join(method_unique_compounds, library, by = c("Compound" = "IA_Name")) %>% select(Compound, DHS, DCSE, Log_P, Vapor_Pressure, Vapor_Pressure_At) -> method_unique_compounds
method_unique_compounds

colSums(method.logical[ , 1:2])
```

A total of `r colSums(method.logical[ , 1:2])[1]` compounds were detected by DHS and `r colSums(method.logical[ , 1:2])[2]` compounds were detected by DCSE.  `r dim(filter(method_unique_compounds, DHS != 0))[1]` of these compounds were unique to DHS sampling and `r dim(filter(method_unique_compounds, DCSE != 0))[1]` were unique to DCSE.

##By treatment
```{r}
data %>% group_by(Treatment) %>% summarise_all(funs(mean)) -> treatment.data
treatment.data[ , -(1:2)] %>% t() %>% as.data.frame() -> treatment.data
colnames(treatment.data) <- c("Control", "MeJA")
treatment.data$Compound <- rownames(treatment.data)
rownames(treatment.data) <- NULL

##Venn Diagram
treatment.logical <- treatment.data
treatment.logical$Control[treatment.logical$Control != 0] <- 1
treatment.logical$MeJA[treatment.logical$MeJA !=0] <- 1

venn(treatment.logical[ , 1:2], cexil = 1, cexsn = .85)

#Extract Unique Compounds
select(treatment.data, Compound, everything()) %>% filter(Control == 0 & MeJA != 0 | Control !=0 & MeJA ==0) -> treatment_unique_compounds
#left_join(treatment_unique_compounds, library, by = c("Compound" = "IA_Name")) %>% select(Compound, Control, MeJA, Log_P, Vapor_Pressure, Vapor_Pressure_At) -> treatment_unique_compounds
treatment_unique_compounds
colSums(treatment.logical[ , 1:2])
```

`r colSums(treatment.logical[ , 1:2])[1]` compounds in Control, ``r colSums(treatment.logical[ , 1:2])[2]`` found in MeJA treated plants.  `r dim(filter(treatment_unique_compounds, Control != 0))[1]` compounds are unique to Control plants and `r dim(filter(treatment_unique_compounds, MeJA != 0))[1]` compounds are unique to MeJA treated plants.


##By treatment AND method
```{r}
data %>% group_by(Treatment, Method) %>% summarise_all(funs(mean)) -> combo.data
combo.data[ , -(1:2)] %>% t() %>% as.data.frame() -> combo.data1
colnames(combo.data1) <- paste(combo.data$Method, combo.data$Treatment, sep = ".")
combo.data1$Compound <- rownames(combo.data1)
rownames(combo.data1) <- NULL

##Venn Diagram
combo.logical <- combo.data1
combo.logical$DHS.Cont[combo.logical$DHS.Cont != 0] <- 1
combo.logical$DCSE.Cont[combo.logical$DCSE.Cont != 0] <- 1
combo.logical$DHS.MeJA[combo.logical$DHS.MeJA != 0] <- 1
combo.logical$DCSE.MeJA[combo.logical$DCSE.MeJA != 0] <- 1

venn(combo.logical[ , 1:4], cexil = 1, cexsn = .85)
colSums(combo.logical[ , 1:4])

#unique to MeJA/DCSE combo
combo.logical %>% filter(DCSE.MeJA == 1, DCSE.Cont == 0, DHS.Cont == 0, DHS.MeJA == 0)

#unique to MeJA/DHS combo
combo.logical %>% filter(DCSE.MeJA == 0, DCSE.Cont == 0, DHS.Cont == 0, DHS.MeJA == 1)

```

## Write to a file
```{r}
#xlsx package broken currently
#write.xlsx(method_unique_compounds, "unique compounds DHS comparison.xlsx", 
#           sheetName = "By Method", row.names = FALSE)
#write.xlsx(treatment_unique_compounds, "unique compounds DHS comparison.xlsx", 
#           sheetName = "By Treatment", row.names = FALSE, append = TRUE)
```

#PCA
I conduct a PCA and plot PC1 vs PC2
```{r}
data.scaled <- scale(data[3:length(data)])
all.pca <- prcomp(data.scaled, scale. = FALSE, center = FALSE)
summary(all.pca)
#4 PCs to capture ~80% of variance
#screeplot(all.pca,type="l")
```
80% of variation captured in ~4 PCs

## PCA Plot
```{r plot theme}
##my ggplot2 theme
multivar.theme <- theme_bw(base_size = 7, base_family = "Helvetica")+  #this changes the default text size and font
  theme(axis.title = element_text(size = 7),                      #this sets a different font size for the axis labels
        panel.border = element_rect(size = .5, color = "black"),   #change size= here to change border size
        panel.grid.major = element_blank(),                     #these two lines make the grid lines go away
        panel.grid.minor = element_blank(),
        legend.text = element_text(size = 6),                  #legend font size
        legend.background = element_rect(color = "black", size = 0.25))
        #legend.key.size = unit(.68, "lines"),
        #legend.spacing = unit(.4, "lines"))         #changes horizontal spacing
```

```{r}
Method <- data$Method
Treatment <- data$Treatment
levels(Treatment) <- c("Control", "MeJA")
Combined <- as.factor(paste(Method, Treatment))

#black and white with open vs. closed triangles and circles
#shape.labels = c("DCSE", "DHS")
#col.labels = c("Control", "MeJA")
#cols <- c("Control" = "NA", "MeJA" = "black")
#ggplot(as.data.frame(all.pca$x), aes(PC1, PC2)) + 
#  geom_point(aes(shape = Combined), size = I(2)) + 
#  xlim(-18, 18) +
#  multivar.theme + 
#  #scale_fill_manual(name = "Method and Treatment",
#                    labels = Combined,
#                    values = cols) +
#  scale_shape_manual(name = NULL,
#                     values = c(16, 17, 1, 2), 
#                     labels = levels(Combined)) +
#  #guides(fill = guide_legend(override.aes = list(color = cols))) +
#  #theme(legend.position = "right", legend.box = "vertical") +
#  ylab(paste0("PC2 (",round(summary(all.pca)$importance[2,2]*100, 2), "%)")) + 
#  xlab(paste0("PC1 (",round(summary(all.pca)$importance[2,1]*100, 2), "%)"))

#ggsave("Comparison PCA bw.png", width = 5, height = 3.33, units = "in")


#ok, for some reason Robbat wants shape to represent method and filled/unfilled to represent treatment

ggplot(as.data.frame(all.pca$x), aes(PC1, PC2)) + 
  geom_point(aes(shape = Combined), size = I(2)) + 
  xlim(-18, 18) +
  multivar.theme + 
  scale_shape_manual(name = NULL,
                     values = c(1, 16, 2, 17), 
                     labels = levels(Combined),
                     guide = guide_legend(nrow = 1)) +
  theme(legend.position = "bottom") +
  theme(legend.key.size = unit(.68, "lines")) +
  theme(legend.margin = margin(t = 2, b = 2, r = 4, l = 1)) +
  ylab(paste0("PC2 (",round(summary(all.pca)$importance[2,2]*100, 2), "%)")) + 
  xlab(paste0("PC1 (",round(summary(all.pca)$importance[2,1]*100, 2), "%)"))
ggsave("Comparison PCA bw4.png", height = 3.5, width = 3.33, units = "in")

```


#Correlations
Need to figure out which variables (chemicals) correlate with PC1 and PC2.  I get this using a function I wrote in myfunctions.R, but I also add on columns for r^2 values and FDR corrected p-values for the correlation between RPA and PC score for each PC.



```{r}
pca.correlations<-PC.cors(all.pca, as.data.frame(data.scaled), 2)
pca.correlations$PC1.r2 <- pca.correlations$PC1.corr.coefs^2
pca.correlations$PC2.r2 <- pca.correlations$PC2.corr.coefs^2
pca.correlations$PC1.p.corrected <- pca.correlations$PC1.p.values %>% p.adjust(method="fdr")
pca.correlations$PC2.p.corrected <- pca.correlations$PC2.p.values %>% p.adjust(method="fdr")

PC1.corr.sig <- pca.correlations %>% subset(PC1.p.corrected <= 0.05 & PC1.r2 >= 0.8, 
                                            select = c(.Variables, PC1.corr.coefs, PC1.r2,
                                                       PC1.p.corrected))
PC2.corr.sig <- pca.correlations %>% subset(PC2.p.corrected <= 0.05 & PC2.r2 >= 0.8,
                                            select = c(.Variables, PC2.corr.coefs, PC2.r2,
                                                       PC2.p.corrected))
PC1.corr.sig[,-1] <- PC1.corr.sig[,-1] %>% signif(digits=3)

PC2.corr.sig[,-1] <- PC2.corr.sig[,-1] %>% signif(digits=3)


#get vapor pressures for these babies

#left_join(PC1.corr.sig, library, by = c(".Variables" = "IA_Name")) %>% select(Compound = .Variables, PC1.corr.coefs, PC1.r2, PC1.p.corrected, Log_P, Vapor_Pressure, Vapor_Pressure_At) -> PC1.corr.sig
arrange(PC1.corr.sig, PC1.corr.coefs)

#left_join(PC2.corr.sig, library, by = c(".Variables" = "IA_Name")) %>% select(Compound = .Variables, PC2.corr.coefs, PC2.r2, PC2.p.corrected, Log_P, Vapor_Pressure, Vapor_Pressure_At) -> PC2.corr.sig
arrange(PC2.corr.sig, PC2.corr.coefs)

#write.xlsx(PC1.corr.sig, "PCA compounds DHS comparison.xlsx", 
#           sheetName = "PC1", col.names = TRUE, row.names = FALSE)
#write.xlsx(PC2.corr.sig, "PCA compounds DHS comparison.xlsx", 
#           sheetName = "PC2", col.names = TRUE, row.names = FALSE, append = TRUE)
write.csv(PC1.corr.sig, "PC1 compounds DHS comparison.csv")
write.csv(PC2.corr.sig, "PC2 compounds DHS comparison.csv")
```

An alternative method might be to use FDR corrected t-tests since it is very clear that PC1 separates the methods and PC2 separates the treatments

```{r}

important.PC1.data <- select(data, one_of(as.character(PC1.corr.sig$.Variables)))

sig.tests.PC1 <- vector(mode = "numeric", length = length(important.PC1.data))
for(i in 1:length(important.PC1.data)){
  test <- t.test(important.PC1.data[[i]] ~ data$Method)
  sig.tests.PC1[i] <- test$p.value
}

important.PC2.data <- select(data, one_of(as.character(PC2.corr.sig$.Variables)))

sig.tests.PC2 <- vector(mode = "numeric", length = length(important.PC2.data))
for(i in 1:length(important.PC2.data)){
  test <- t.test(important.PC2.data[[i]] ~ data$Treatment)
  sig.tests.PC2[i] <- test$p.value
}
```
The only thing this changes is it takes Benzyl Alcohol off the list for PC2

#ANOVA (really just t-tests?)
Using the PCA scores, test if there is significant difference between methods, treatments, interaction?? (2 way ANOVA)
```{r}
df <- data.frame(treatment = Treatment, method = Method, PC1 = all.pca$x[ , 1], PC2 = all.pca$x[ , 2])
df
#interaction was non-significant
anova(lm(PC1 ~ treatment+method, data = df))
t.test(PC1 ~ method, data = df)

anova(lm(PC2 ~ treatment+method, data = df))
t.test(PC2 ~ treatment, data = df)
```
