---
title: "Proof of concept Field Experiment"
output: html_notebook
---

This is code migrated over from TwisterPCA.R and tidied up
```{r}
setwd("/Users/scottericr/Google Drive/R Code/Twister Data")
#setwd("/Users/escott03/Google Drive/R Code/Twister Data")
library(reshape2)
library(dplyr)
library(ggplot2)
library(readr)
library(xlsx)
#library(car)
#library(scales)
#library(MASS)
library(RColorBrewer)
library(venn)
source("/Users/scottericr/Google Drive/R Code/myfunctions.R")
```

read in data (updated to Field2.xlsx on Mar 30, 2017)
```{r}
#twister <- read_csv("Field.csv")
twister <- read.xlsx("Field2.xlsx", sheetIndex = 3)
twister1 <- twister[1:149, 1:12]
twister1[is.na(twister1)] = 0 #missing cells represent undetected compounds
twist2 = recast(twister1, variable ~ Compound, measure.var = c(2:12)) #this transposes the data so each column is a compound
twist2
colnames(twist2)[1] <- "samples"
#glimpse(twist2)
summary(twist2$samples)

Treatment<-c(rep("E. obliqua",3),rep("MeJA",4),rep("control",4))
```
#Exploratory stats
how many compounds detected total?

```{r}
dim(twist2)-1 #149
```
how many compounds detected in each treatment?

```{r}
x <- transmute(twister1, Control = con1 + con2 + con3 + con4,
               MeJA = meja1 + meja2 + meja3 + meja4,
               Geometrid = geo1 + geo2 + geo3)
x[x > 0] = 1
x$Compound <- colnames(twist2)[-1]
x <- dplyr::select(x, Compound, everything())
venn(x[ , -1], cexil=1.5, cexsn=1.1)

colSums(x[ , -1])
#unique to Geometrid: 
Geometrid.unique <- subset(x, x$Geometrid != 0 & x$MeJA == 0 & x$Control == 0)
Geometrid.unique

#unique to MeJA:
MeJA.unique <- subset(x, x$MeJA != 0 & x$Geometrid == 0 & x$Control == 0)
MeJA.unique

#unique to Control
Control.unique <- subset(x, x$Control != 0 & x$Geometrid == 0 & x$MeJA == 0)
Control.unique
```
# PCA
```{r}
tea.pca2<-prcomp(twist2[ , -1], scale. = TRUE, center = TRUE)
summary(tea.pca2)
#head(tea.pca2$rotation)
#screeplot(tea.pca2, type="lines")
```

```{r plot theme}
##my ggplot2 theme
multivar.theme <- theme_bw(base_size = 10, base_family = "Arial")+  #this changes the default text size and font
  theme(axis.title = element_text(size = 10),                      #this sets a different font size for the axis labels
        panel.border = element_rect(size = 1, color = "black"),   #change size= here to change border size
        panel.grid.major = element_blank(),                     #these two lines make the grid lines go away
        panel.grid.minor = element_blank())

```

```{r, fig.height=2.75, fig.width=3.5}
####Plotting PCA####
#Treatment

labels = expression("Control", italic("E. obliqua"), "MeJA")
cols <- c("control" = "#1b9e77", "MeJA" = "#d95f02", "E. obliqua" = "#7570b3")


#PC1 vs PC2
ggplot(as.data.frame(tea.pca2$x), aes(PC1, PC2, fill = Treatment)) + 
  geom_point(shape = 24, size = I(3)) + 
  scale_fill_manual(values = cols, labels = labels) +
  multivar.theme + 
  theme(
    legend.position = "right",
  #legend.position=c(0.13,0.16), #puts it in the lower left
    legend.text = element_text(size = 10),
    legend.background = element_rect(color = "black")) +
  ylab(paste0("PC2 (",round(summary(tea.pca2)$importance[2,2]*100, 2), "%)")) + 
  xlab(paste0("PC1 (",round(summary(tea.pca2)$importance[2,1]*100, 2), "%)"))
#ggsave("Field Trial.png", height = 3.5, width = 5, units = "in")


#PC2 vs PC3
#ggplot(as.data.frame(tea.pca2$x), aes(PC2, PC3, fill = Treatment)) + 
#  geom_point(shape = 24, size = I(3)) + 
#  scale_fill_manual(values = cols, labels = labels) +
#  multivar.theme + 
#  theme(
#    legend.position = "right",
#  #legend.position=c(0.13,0.16), #puts it in the lower left
#    legend.text = element_text(size = 10),
#    legend.background = element_rect(color = "black")) +
#  ylab(paste0("PC3 (",round(summary(tea.pca2)$importance[2,3]*100, 2), "%)")) + 
#  xlab(paste0("PC2 (",round(summary(tea.pca2)$importance[2,2]*100, 2), "%)"))
#ggsave("Field Trial2vs3.png", height = 3.5, width = 5, units = "in")
```

```{r}
pca.correlations <- PC.cors(tea.pca2, as.data.frame(twist2[ , -1]), 2)
pca.correlations$PC1.r2 <- pca.correlations$PC1.corr.coefs^2
pca.correlations$PC2.r2 <- pca.correlations$PC2.corr.coefs^2
pca.correlations$PC1.p.corrected <- pca.correlations$PC1.p.values %>% p.adjust(method="fdr")
pca.correlations$PC2.p.corrected <- pca.correlations$PC2.p.values %>% p.adjust(method="fdr")

PC1.corr.sig <- pca.correlations %>% subset(PC1.p.corrected <= 0.05 & PC1.r2 >= 0.7, 
                                            select = c(.Variables, PC1.corr.coefs, PC1.r2,
                                                       PC1.p.corrected))
PC2.corr.sig <- pca.correlations %>% subset(PC2.p.corrected <= 0.05 & PC2.r2 >= 0.7,
                                            select = c(.Variables, PC2.corr.coefs, PC2.r2,
                                                       PC2.p.corrected))
PC1.corr.sig[,-1] <- PC1.corr.sig[,-1] %>% signif(digits=3)
PC1.corr.sig
PC2.corr.sig[,-1] <- PC2.corr.sig[,-1] %>% signif(digits=3)
PC2.corr.sig

write.xlsx(PC1.corr.sig, "Field Trial PCA compounds.xlsx", sheetName = "PC1", col.names = TRUE, row.names = FALSE)
write.xlsx(PC2.corr.sig, "Field Trial PCA compounds.xlsx", sheetName = "PC2", col.names = TRUE, row.names = FALSE, append = TRUE)
```

Nicole asks "Why is `γ-Terpinene` not highly correlated with PC1 when it is absent from control samples but present in both geometrid and MeJA treatments?"

Well, PC1 is the axis that explains the most variation in the data, so it's likely that even though `γ-Terpinene` might be a good predictor of herbivory, it's not super variable. Let's look at what its loading is on PC1

```{r}
tea.pca2$rotation[rownames(tea.pca2$rotation) == "γ-Terpinene" , 1]
head(sort(tea.pca2$rotation[ ,1]))
#which PC has the greatest loading for gamma terpinene?
tea.pca2$rotation[rownames(tea.pca2$rotation) == "γ-Terpinene", ]
#PC3
head(sort(tea.pca2$rotation[ , 3]))
```
It's not very heavily weighting PC1 compared to other compounds.  Maybe it is at very low concentrations when it is present.  A bunch of zeros and a bunch of low concentration values would have very low variance.

```{r}
select(twist2, samples, `γ-Terpinene`)
```


An alternative method might be to use FDR corrected t-tests since it is very clear that PC1 separates the methods and PC2 separates the treatments

```{r}

important.PC1.data <- select(data, one_of(as.character(PC1.corr.sig$.Variables)))

sig.tests.PC1 <- vector(mode = "numeric", length = length(important.PC1.data))
for(i in 1:length(important.PC1.data)){
  test <- t.test(important.PC1.data[[i]] ~ data$Method)
  sig.tests.PC1[i] <- test$p.value
}

important.PC2.data <- select(data, one_of(as.character(PC2.corr.sig$.Variables)))

sig.tests.PC2 <- vector(mode = "numeric", length = length(important.PC2.data))
for(i in 1:length(important.PC2.data)){
  test <- t.test(important.PC2.data[[i]] ~ data$Treatment)
  sig.tests.PC2[i] <- test$p.value
}
```
